---
title: "Analysis Of Temperatures Data from Prague Klementinum Weather Station"
author: "Jan Schubert"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 10, warning = FALSE, message = FALSE)
```

```{r package and function load}
library(data.table)
library(rvest)
library(strucchange)
library(highcharter)
library(xts)

## SCRAPE
scrapeWeather <- function(weatherUrl) {
  #Reading the HTML code from the website
  webpage <- read_html(weatherUrl)
  #Using CSS selectors to scrap the title section
  weather_data_html <- html_nodes(webpage, '.prehled')
  #Converting the title data to text
  weather_data <-
    html_table(weather_data_html, fill = TRUE, header = TRUE)
  #format to a data.frame
  weather_data <- as.data.frame(weather_data[[1]])
  #loose the NA rows
  weather_data <- weather_data[!is.na(weather_data$Stanice),]
  return(weather_data)
}
```

```{r data scape, eval=FALSE}
# url
seqDates <- seq.Date(from = as.Date("1775-01-01"), 
                     to = Sys.Date(), 
                     by = "day")
seqDates_url <- paste(format(seqDates, format = "%d"), format(seqDates, format = "%m"), format(seqDates, format = "%Y"), sep = "-")

weather_list <- rep(list(NULL),length(seqDates))
names(weather_list) <- seqDates
for(i in seq_along(seqDates_url)){
  tryCatch({
    weatherUrl <- paste0("https://www.in-pocasi.cz/archiv/archiv.php?historie=",seqDates_url[i],"&stanice_kraj=0&klima_kraj=0#klima")
  weather_list[[i]] <- scrapeWeather(weatherUrl)
  },  error=function(e){})
}
```

```{r load chached data}
load("weatherData.Rdata")

weatherDat <- lapply(weather_list, function(x) {
  x[x$Stanice == "Praha - Klementinum", ]
})
weatherDat <- as.data.table(do.call(rbind, weatherDat),keep.rownames = TRUE)
names(weatherDat)[c(1,3:8)] <- c("Day", "Max.Temperature", "Min.Temperature", "Wind", "Precipitation", "Max.Pressure", "Max.Humidity")
selCol <- names(weatherDat)[3:8]
weatherDat[, selCol[1:2] := list(gsub(" °C","",Max.Temperature), gsub(" °C","",Min.Temperature))]
weatherDat[, selCol[3] := gsub(" km/h","",Wind)]
weatherDat[, selCol[4] := gsub(" mm","",Precipitation)]
weatherDat[, selCol[5] := gsub(" hPa","",Max.Pressure)]
weatherDat[, selCol[6] := gsub("%","",Max.Humidity)]
weatherDat[, c(selCol) := lapply(.SD, as.numeric), .SDcols = selCol]
#format date
weatherDat[, Day := as.Date(Day, format = "%Y-%m-%d")]
#weather_list <- NULL
```


## Long Term Trend In Average Yearly Temperature
We use data from a weather station in Prague Klementinum. The start of the data is in January 1775, the end in December 2017.

The chart below shows the average yearly temperature and a smoothed trend. The temperature has been wobbling between 12 and 13C till the beginning of the 20th century at which point it raises above 13C and stabilizes. It has risen exponentially since the 80s and hit the 14C and 15C marks.

```{r plot yearly}
weatherDat.yearly <- weatherDat[, .(Max.Temperature = mean(Max.Temperature)), 
                                 by = .(Year = format(Day, format = "%Y"))]
weatherDat.yearly[, Year := as.Date(paste0(Year,"-01-01"), format = "%Y-%m-%d")]
m_loess <- loess(weatherDat.yearly$Max.Temperature ~ seq_along(weatherDat.yearly$Max.Temperature), span = .3)
Max.Temperature_trend <- predict(m_loess)

Max.Temperature <- ts(data = weatherDat.yearly$Max.Temperature, 
                            start = as.numeric(format(head(weatherDat.yearly$Year,1), format = "%Y")), 
                            end = as.numeric(format(tail(weatherDat.yearly$Year,1), format = "%Y")), 
                            frequency = 1)
Max.Temperature_trend <- ts(data = Max.Temperature_trend, 
                            start = as.numeric(format(head(weatherDat.yearly$Year,1), format = "%Y")), 
                            end = as.numeric(format(tail(weatherDat.yearly$Year,1), format = "%Y")), 
                            frequency = 1)

hchart(as.xts(Max.Temperature)) %>%
  hc_add_series(as.xts(Max.Temperature_trend))
```

## Structural Changes in Average Yearly Temperature
In this section, we want to test if the recent rise in the average yearly temperature is part of a random process or if (given the data) this increase is unlikely to be due to random changes in the temperature. The statistical test we use looks at significant changes in average temperature over all possible time frames in our time series. We define the length of one time frame to be 20 years which gives us enough stability to capture longer trends in temperature.


Using this analysis, we found 4 segments in our temperature data:

1. 1775 till 1836 with an average temperature of 13C

2. 1837 till 1924 with an average temperature of 12.5C

3. 1925 till 1987 with an average temperature of 13.6C

4. 1988 till 2017 with an average temperature of 15C.

There are two interesting observations. One is the obvious sudden increase in the last 30 years (1987 till 2017). The second is the increasing confidence in where a breakpoint is (shown by the red line at the bottom of the chart). Whilst the first breakpoint is likely to happen between the years 1809 to 1836 (an interval of 27 years), the last breakpoint is likely between the years 1983 and 1987, just 4 years. This reflects the decreasing randomness in the temperatures supporting a hypothesis that the temperature in recent years has been systematically increasing.

```{r}
temperature_brp <- breakpoints(Max.Temperature ~ 1, h = 20)

plot(Max.Temperature, main = "Average Yearly Temprerature - Praha Klementinum")
lines(fitted(temperature_brp, breaks = length(temperature_brp$breakpoints)), col = 4)
lines(confint(temperature_brp, breaks = length(temperature_brp$breakpoints)))
```


## Distribution of Yearly Temperatures
The chart below shows the distribution of yearly temperatures. The lines below show the individual years. Lines in red indicate years from the last segment (1988 - 2017). Most of the years are the hottest years ever recorded. In fact, all of the top 10 hottest years are from this era (and 18 of the 20 hottest years).

```{r}
temperature_seg1 <- Max.Temperature[1:temperature_brp$breakpoints[3]]
temperature_seg2 <- Max.Temperature[(temperature_brp$breakpoints[3]+1):length(Max.Temperature)]

hist(Max.Temperature, main = "Distribution Of Yearly Temperatures")
rug(jitter(temperature_seg1, amount = 0.01), col = "black")
rug(jitter(temperature_seg2, amount = 0.01), col = "red", lwd = 3)
legend("topright", legend = c("1775 - 1987", "1988 - 2017"), col = c("black", "red"), lty = c(1,1))

head(weatherDat.yearly[order(Max.Temperature, decreasing = TRUE), ], 20)
```